---
title: "Cleaning of the Marijuana Prices Dataset"
author: "Alejandro Romero González"
affiliation: "Banco de México"
date: "21/02/2026"
---

```{r}
library(here)
library(dplyr)
library(stringr)
library(lubridate)
library(arrow)
```

First, we load the raw dataset:
```{r}
RawData <- read.csv(here("Data", "Marijuana_Prices_in_Mexico_raw-3_utf8_bom.csv"))
```

The dataset contains the following structure:
```{r}
glimpse(RawData)
```

Second, we standarize variable names, build new columns, drop unnecessary variables, and filter out invalid observations:
```{r}
RawData <- RawData |>
  rename(
    response_id = ResponseID,
    start_date = StartDate,
    end_date = EndDate
  ) |>
  select(-response_id, -experience, -frequency, -average_amount, -relationship, 
         -status, -start_date, -end_date) |>
  mutate(
    quality = str_to_lower(str_trim(quality)),
    state = str_squish(state),
    municipality = str_squish(municipality),
    gender = str_squish(gender),
    education = str_squish(education),
    
    # unify amount/value across good vs bad columns
    quantity_g = coalesce(good_quantity, bad_quantity),
    value_mxn  = coalesce(good_value,  bad_value),
    price_mxn_g = value_mxn / quantity_g
  ) |>
  select(-good_quantity, -bad_quantity, -good_value, -bad_value)

RawData <- RawData |>
  filter(
    !is.na(quality),
    !is.na(state),
    !is.na(quantity_g),
    !is.na(value_mxn),
    
    age<99,
    value_mxn > 0,
    value_mxn < 9999999,
    quantity_g >= 1,
    
    municipality!="",
    purchase_date!="",
    gender!="",
    education!="",
    state != ""
  ) |>
  mutate(
    quality = factor(quality, levels = c("bad", "good")),
    
    # Decompose municipality into code and name
    muni_code = as.integer(str_extract(municipality, "\\d{5}$")),
    muni_name = str_trim(str_remove(municipality, "\\s*\\d{5}$"))
  ) |>
  select(-municipality)
```

Next, we have to solve an issue with the _purchase_date_ varible, which currently has two different formats:
```{r}
RawData <- RawData |>
  mutate(
    purchase_date = parse_date_time(
      purchase_date,
      orders = c("dby", "d-b-y", "d-b-Y")
    ),
    purchase_date = as.Date(purchase_date)
  ) |>
  filter(purchase_date >= as.Date("2019-01-01"))
```

We drop purchases dated before 2019-01-01. Although the survey ran from February 19 to September 15, 2019, some respondents reported purchases from earlier years. Restricting to 2019 purchases improves temporal comparability and reduces the risk that recalled transactions from prior years reflect different market conditions, price levels, or reporting noise. After all filters, the analytic sample contains 3,293 observations dated January 1 to September 11, 2019.

**Summary of sample restrictions** (applied sequentially from 4,775 raw records):

| Filter | Obs dropped | Rationale |
|--------|------------|-----------|
| Missing required fields | 1,291 | Missing transaction, geography, date, or demographic fields |
| Implausible age, value, or quantity | 7 | Implausible placeholder entries, data entry errors, or likely unit confusion |
| Purchase date before 2019-01-01 | 184 | Recalled transactions from prior years |
| **Final analytic sample** | — | **N = 3,293** |

Note: filter counts are sequential after standardizing variables and coalescing the quality-specific quantity and expenditure fields.

Finally, we reorder the columns and save the cleaned dataset:
```{r}
CleanData <- RawData |>
  select(gender, age, education, state, muni_code, muni_name, 
         purchase_date, quantity_g, quality, value_mxn, price_mxn_g)

write_parquet(CleanData, here("Data", "Marijuana_Prices_in_Mexico_clean.parquet"))
```
